### 2024-04-18 manuscript.qmd
first commit

---
title: "Machine Learning Approaches for Ammonia Volatilization Prediction after Manure Application"
author: 
  - "Armand Favrot"
  - "Sophie Génermont"
  - "Céline Decuq"
  - "Jean-Marc Giliot"
  - "David Makowski"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library ("ggplot2") 
library ("gridExtra")
library ("dplyr")       
library ("paletteer")       
library ("tibble") # rownames_to_column
library ("tidyr") # pivot_longer
library ("ALFAM2")
library ("randomForest")
library ("xgboost")
library ("glmnet")
library ("shapviz")
library ("treeshap")

set.seed (123)
```

```{r}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

theme_replace(
    axis.text = element_text(size = axis_text_size), 
    axis.title.x = element_text(size = axis_title_size, angle = 0, margin = ggplot2::margin (t = 10)), 
    axis.title.y = element_text(size = axis_title_size, angle = 90, margin = ggplot2::margin (r = 10)), 
    plot.title = element_text(size = title_size, hjust = 0.5, face = "bold"), 
    legend.text = element_text(size = axis_title_size), 
    legend.title = element_text(size = axis_title_size, face = "bold"), 
    strip.text = element_text(size = axis_title_size, face = "bold", color = "white"), 
    strip.background = element_rect(fill = "#000080"), 
    panel.background = element_rect(fill = "#f3faff"), panel.grid.major = element_line(colour = "white"), 
    panel.spacing = unit(2, "lines")
)

options(ggplot2.discrete.fill = list("#67c5a7"))

Dark2 = paletteer_d("RColorBrewer::Dark2")
```

# Data description

```{r}
load (file = "scripts/data_alfam2.Rdata")
```

```{r}
plot = data_alfam2 %>%

    filter (! (app.mthd == "bsth" & incorp == "shallow")) %>%

    mutate (strategy = paste (app.mthd, incorp, sep = " - ")) %>%
    mutate (strategy = recode (strategy, "bc - none" = "A", "bc - shallow" = "B", "bc - deep" = "C", "bsth - none" = "D", "os - none" = "E", "ts - none" = "F")) %>%

    ggplot () +
        geom_line (aes (x = time, y = e.rel, group = pmid), linewidth = 0.15) +
        ylab ("Cumul of relative emission") +
        xlab ("Time (h)") +
        theme (axis.title.y = element_text (margin = ggplot2::margin (r = 25))) +
        # scale_color_manual (values = Dark2) +
        facet_wrap (~ strategy)

png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
plot
dev.off()

plot
```

# Part 1 - Model comparison

## Alfam2

```{r}
pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]

alfam2_predictions =  alfam2 (

        # Important to use alfam2pars01 parameters, which correspond to the parameters used in the paper
        # New parameters are available, but leads to ohter results (less good results)
        pars = pars,

        dat = data_alfam2 %>% select (- j.NH3, - e.cum, - e.rel, - dataset),

        app.name = "tan.app",

        time.name = "time",

        time.incorp = "t.incorp",

        group = "pmid",

        prep = TRUE,

        warn = FALSE
)
```

We add the true values to the prediction dataframe and we keep only the last observation of each pmid.

```{r}
alfam2_predictions = alfam2_predictions %>%

        select (pmid, time, e, er) %>%

        mutate (truth_e = data_alfam2$e.cum) %>%
        mutate (truth_er = data_alfam2$e.rel) %>%

        mutate (dataset = data_alfam2$dataset) %>%

        select (pmid, time, e, er, truth_e, truth_er, dataset) %>%

        filter (time == max (time), .by = pmid)
```

## Random forest

```{r}
load (file = "scripts/data_random_forest.Rdata")
```

```{r}
data_random_forest_calibration = data_random_forest %>% 
    filter (dataset == "Calibration subset") %>%
    select (- pmid, - e.rel, - dataset)

random_forest_model = randomForest (

        e.cum ~ .,

        data = data_random_forest_calibration,

        importance = TRUE,

        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8

)
```

```{r}
plot (random_forest_model)
```

```{r}
random_forest_predictions = data_random_forest %>%

    mutate (e.cum_hat = predict ( 
                random_forest_model,                       
                newdata = data_random_forest %>% 
                    select (- pmid, - e.cum, - e.rel, - dataset)
            )
    ) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)    
```

## Xgboost

```{r}
load (file = "scripts/data_xgboost.Rdata")
```

```{r}

xgboost_model = xgboost (

    data = xgb.DMatrix (

        data = data_xgboost %>% 
            filter (dataset == "Calibration subset") %>%
            select (- pmid, - e.cum, - e.rel, - dataset) %>%
            as.matrix %>%
            {.},

        label = data_xgboost %>% 
            filter (dataset == "Calibration subset") %>%
            select (e.cum) %>% 
            as.matrix %>%
            {.}

    ),

    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,

    verbose = FALSE,

    objective = "reg:squarederror"

)
```

```{r}
xgboost_predictions = data_xgboost %>%

    mutate (e.cum_hat = predict (

                xgboost_model, 
                newdata = data_xgboost %>% 
                        select (- pmid, - e.cum, - e.rel, - dataset) %>% 
                        as.matrix
            )
    ) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)
```

## Lasso

```{r}
load (file = "scripts/data_lasso.Rdata")
```

```{r}
lasso_model = cv.glmnet (

    x = data_lasso %>% 
        filter (dataset == "Calibration subset") %>%
        select (- pmid, - e.cum, - e.rel, - dataset) %>%
        as.matrix %>%
        {.},

    y = data_lasso %>% 
        filter (dataset == "Calibration subset") %>%
        select (e.cum) %>% 
        as.matrix %>%
        {.},

    alpha = 1
)
```

```{r}
lasso_predictions_vector = predict (

    lasso_model,

    data_lasso %>% 
            select (- pmid, - e.cum, - e.rel, - dataset) %>% 
            as.matrix
)

lasso_predictions = data_lasso %>%

    mutate (e.cum_hat = exp (lasso_predictions_vector), e.cum = exp (e.cum)) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset) 
```

## Comparison

```{r}
predictions_of_all_methods = rbind (
    alfam2_predictions %>% mutate (method = "alfam2"), 
    random_forest_predictions %>% mutate (method = "random forest"),
    xgboost_predictions %>% mutate (method = "xgboost"),
    lasso_predictions %>% mutate (method = "lasso")
)
```

### Observed vs predicted values

```{r}
plot = predictions_of_all_methods %>%

    filter (dataset == "Evaluation subset") %>%

    mutate (method = recode (method, "alfam2" = "A", "lasso" = "B", "random forest" = "C", "xgboost" = "D")) %>%

    ggplot () +

        geom_point (aes (x = truth_e, y = e, color = method), size = 3) +

        geom_abline (slope = 1, linetype = "dashed") +

        facet_wrap (~ method) +

        scale_color_manual (values = Dark2[c(2, 5, 6, 8)]) +

        labs (color = "") +

        theme (legend.position = "none",
               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
               axis.title.x = element_text (margin = ggplot2::margin (t = 15, b = 20))) +

        xlab ("Observed values") + ylab ("Predicted values") +

        NULL

png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
plot
dev.off()

plot
```

## Evaluation metrics

```{r}
df_evaluation_metrics = rbind (

    predictions_of_all_methods %>% 
        select (prediction = e, truth = truth_e, dataset, method) %>% 
        mutate (name = "72h cum. emission"),

    predictions_of_all_methods %>% 
        select (prediction = er, truth = truth_er, dataset, method) %>% 
        mutate (name = "72h relative cum. emission")

) %>%

    summarise (

        Pearsons_r = cor (prediction, truth),

        ME = 1 - (sum ( (prediction - truth) ^ 2) / sum ( (truth - mean (truth)) ^ 2)),

        MAE = mean (abs (prediction - truth)),

        MBE = mean (prediction - truth),

        .by = c (name, dataset, method)

) %>%

    mutate_if (is.numeric, round, digits = 3)

df_evaluation_metrics
```

```{r}
df_evaluation_metrics %>%

    rename (response = name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE"))) %>% 
    filter (name %in% c ("MAE", "MBE")) %>%

    arrange (name, dataset) %>%

    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(response, name, dataset))
```

```{r}
# Evaluation for 72h cum. emission --------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "none ", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------

# Evaluation for 72h relative cum. emission ----------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "bottom", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------
```

# Part 2 - New training with the whole dataset

## Random forest

```{r}
random_forest_model = randomForest (

        e.cum ~ .,

        data = data_random_forest %>% select (- pmid, - e.rel, - dataset),

        importance = TRUE,

        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8

)
```

### Shapley

```{r}
unified_model_rf <- randomForest.unify(random_forest_model, data_random_forest_calibration %>% select (- e.cum))

treeshap_rf <- treeshap(unified_model_rf, data_random_forest_calibration %>% select (- e.cum))

shapley_rf <- shapviz(treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
```

```{r}
plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
    scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
    theme (
      legend.title = element_text (size = 25, margin = ggplot2::margin (r = 50, b = 100)),
      legend.position = "bottom",
      legend.key.width = unit(3, "cm"),
      axis.title.x = element_text (margin = ggplot2::margin( b = 30))
    )

plot_shap_rf
```

### Gain

```{r}
df_importance_rf = importance (random_forest_model) %>%
    as.data.frame %>%
    rownames_to_column (var = "Variable") %>%
    mutate (`%IncMSE` = `%IncMSE` / max (`%IncMSE`)) %>%
    mutate (type = "A")

plot_importance_rf = ggplot (df_importance_rf) + 
        geom_point (aes (x = `%IncMSE`, y = Variable), size = 5) +
        scale_y_discrete (limits = df_importance_rf %>% arrange (`%IncMSE`) %>% pull (Variable)) +
        xlab ("Importance") +
        ylab ("") +
        facet_wrap (~ type)

plot_importance_rf
```
## Xgboost

### Shapley

```{r}
xgboost_model = xgboost (

    data = xgb.DMatrix (

        data = data_xgboost %>% 
            select (- pmid, - e.cum, - e.rel, - dataset) %>%
            as.matrix %>%
            {.},

        label = data_xgboost %>% 
            select (e.cum) %>% 
            as.matrix %>%
            {.}

    ),

    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,

    verbose = FALSE,

    objective = "reg:squarederror"

)
```

```{r}
df_tmp = data_xgboost %>% select (- pmid, - e.cum, - e.rel, - dataset)

unified_model_xgboost <- xgboost.unify(xgboost_model, as.matrix(df_tmp))

treeshap_xgboost <- treeshap(unified_model_xgboost, df_tmp)

shapley_xgb <- shapviz(treeshap_xgboost, X = df_tmp)
```

```{r}
plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
  scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
  theme (legend.title = element_text (size = 30, margin = ggplot2::margin (r = 50, b = 100)),
         legend.position = "bottom",
         legend.key.width = unit(3, "cm"),
         axis.title.x = element_text (margin = ggplot2::margin( b = 30)))

plot_shap_xgboost
```

### Gain

```{r}
df_importance_xgb = xgb.importance (model = xgboost_model) %>%
    mutate (Gain = Gain / max (Gain)) %>%
    mutate (type = "B")

plot_importance_xgboost = df_importance_xgb %>%

    ggplot() +
        geom_point (aes (x = Gain, y = Feature), size = 5) +
        scale_y_discrete (limits = df_importance_xgb %>% arrange (Gain) %>% pull (Feature)) +
        xlab ("Importance") + ylab ("") +
        facet_wrap (~ type)

plot_importance_xgboost
```
# Part 3 - Testing scenarios

```{r}
load (file = "scripts/data_scenarios.Rdata")

data_scenarios = data_scenarios %>%

    mutate (t.incorp = as.factor (t.incorp)) %>%
    mutate (t.incorp = recode (t.incorp, "1000" = "-"))
```

## Random forest

```{r}
load (file = "scripts/data_scenarios_random_forest.Rdata")

data_tmp = data_scenarios_random_forest

random_forest_predictions_scenarios_vector = predict (
        random_forest_model,       
        newdata = rbind (data_random_forest_calibration[1, -1], data_tmp)[- 1, ]
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_random_forest = random_forest_predictions_scenarios_vector, .before = time) %>%

    mutate (efficacy_random_forest = ((e.cum_hat_random_forest / e.cum_hat_random_forest[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain)) %>%

    {.}
```

## Xgboost

```{r}
load (file = "scripts/data_scenarios_xgboost.Rdata")

xgboost_predictions_scenarios_vector = predict (

    xgboost_model,

    data_scenarios_xgboost %>% as.matrix
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_xgboost = xgboost_predictions_scenarios_vector, .before = time) %>%
    mutate (e.cum_hat_xgboost = ifelse (e.cum_hat_xgboost < 0, 0, e.cum_hat_xgboost)) %>%

    mutate (efficacy_xgboost = ((e.cum_hat_xgboost / e.cum_hat_xgboost[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Lasso

```{r}
load (file = "scripts/data_scenarios_lasso.Rdata")

lasso_predictions_scenarios_vector = predict (

    lasso_model,

    data_scenarios_lasso %>% as.matrix
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_lasso = lasso_predictions_scenarios_vector, .before = time) %>%

    mutate (efficacy_lasso = ((e.cum_hat_lasso / e.cum_hat_lasso[app.mthd == "bc" & incorp == "none"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Alfam2

```{r}
load (file = "scripts/data_scenarios_alfam2.Rdata")

alfam2_predictions_scenarios_vector =  alfam2 (
        pars = alfam2pars01,
        dat = data_scenarios_alfam2,
        app.name = "tan.app",
        time.name = "time",
        time.incorp = "t.incorp",
        group = "pmid",
        prep = TRUE,
        warn = FALSE
)  %>%

    filter (time == max (time), .by = pmid) %>%
    select (pmid, e.cum_hat_alfam2 = e)

data_scenarios = data_scenarios %>%

    left_join (alfam2_predictions_scenarios_vector, by = "pmid") %>%

    mutate (efficacy_alfam2 = ((e.cum_hat_alfam2 / e.cum_hat_alfam2[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Comparison

```{r}
plot = data_scenarios %>% 
    select (app.mthd, incorp, t.incorp, man.ph,
            efficacy_random_forest, efficacy_xgboost, efficacy_lasso, efficacy_alfam2) %>%

    rename (efficacy_rforest = efficacy_random_forest) %>%

    pivot_longer (cols = c (5 : 8), names_to = c (".value", "method"), names_pattern = "(.+)_(.+)") %>%
    mutate (method = recode (method, "rforest" = "random forest")) %>%

    mutate (method = recode (method, "alfam2" = "A", "lasso" = "B", "random forest" = "C", "xgboost" = "D")) %>%

    mutate (group = paste (app.mthd, incorp, t.incorp)) %>%
    filter (! (app.mthd == "bc" & incorp == "none")) %>%

    mutate (group = recode (group,
                            "bc shallow 0" = "incorporation",
                            "bsth none -" = "trailing hose",
                            "os none -" = "open slot",
                            "ts none -" = "trailing shoe")) %>%

    ggplot () +

        geom_boxplot (aes (x = efficacy, y = group, fill = group)) +
        scale_fill_manual (values = Dark2[c(2 : 5)]) +
        ylab ("") + xlab ("Efficacy") +
        theme (legend.position = "none",
               axis.ticks.y = element_blank()) +
        facet_wrap (~ method, ncol = 1)

png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
plot
dev.off()

plot
```

# Figures

```{r}
legend = cowplot::get_legend (plot_shap_rf)

plot_shap_rf_2 = plot_shap_rf +
    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) +
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())

plot_shap_xgboost_2 = plot_shap_xgboost +
    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) + 
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.title.x = element_text (margin = ggplot2::margin (t = 10, b = 20)))

png (file = "figures/plot_shap.png", width = 1000, height = 900)
do.call ("grid.arrange", 
         c (list(plot_shap_rf_2, 
                 plot_shap_xgboost_2, 
                 legend), 
            list(layout_matrix = as.matrix(c (1, 2, 3)),
                  heights = c (0.8, 1, 0.2))))
dev.off()
```

```{r}
png (file = "figures/plot_importance.png", width = 1300, height = 800)
grid.arrange (plot_importance_rf, plot_importance_xgboost, nrow = 1)
dev.off()
```

# Supplementary material

### 2024-04-18 scripts/data_processing.qmd
first commit

---
title: "Data processing"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library ("ggplot2") 
library ("dplyr")       
library ("paletteer")       
library ("tibble") # rownames_to_column
library ("tidyr") # pivot_longer
```

```{r}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

theme_replace(
    axis.text = element_text(size = axis_text_size), 
    axis.title.x = element_text(size = axis_title_size, angle = 0, margin = ggplot2::margin (t = 10)), 
    axis.title.y = element_text(size = axis_title_size, angle = 90, margin = ggplot2::margin (r = 10)), 
    plot.title = element_text(size = title_size, hjust = 0.5, face = "bold"), 
    legend.text = element_text(size = axis_title_size), 
    legend.title = element_text(size = axis_title_size, face = "bold"), 
    strip.text = element_text(size = axis_title_size, face = "bold", color = "white"), 
    strip.background = element_rect(fill = "#000080"), 
    panel.background = element_rect(fill = "#f3faff"), panel.grid.major = element_line(colour = "white"), 
    panel.spacing = unit(2, "lines")
)

options(ggplot2.discrete.fill = list("#67c5a7"))

Dark2 = paletteer_d("RColorBrewer::Dark2")
```

# data_alfam2

```{r}
data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
data_plot_level = read.csv (file = "../data/ALFAM2_plot.csv")

calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
evaluation_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc2.csv")

pmids = rbind (evaluation_subset_pmids %>% mutate (dataset = "Evaluation subset"),
               calibration_subset_pmids %>% mutate (dataset = "Calibration subset"))
```

```{r}
data_alfam2 = data_interval_level %>%

    filter (pmid %in% pmids$pmid) %>%
    left_join (pmids, by = "pmid") %>%

    filter (ct > 0 & ct <= 78) %>%
    select (j.NH3, e.cum, e.rel, pmid, ct, dt, air.temp, wind.2m, rain.rate, rain.cum, dataset)  %>%

    # adding plot level informations --------
    left_join (

        data_plot_level %>%
            filter (pmid %in% pmids$pmid) %>%
            select (pmid, tan.app, app.method, app.rate, man.dm, man.ph, man.source, incorp, time.incorp) %>%
            rename (app.mthd = app.method, t.incorp = time.incorp),

        by = "pmid"

    ) %>% 
    # --------

    replace_na (list (rain.rate = 0)) %>% 
    replace_na (list (rain.cum = 0)) %>%

    mutate (wind.2m = replace_na (wind.2m, mean (wind.2m, na.rm = TRUE)), .by = pmid) %>%
    mutate (air.temp = replace_na (air.temp, mean (air.temp, na.rm = TRUE)), .by = pmid) %>%

    # we keep only trials that last more than 24h, to be in the same settings as for the machine learning models
    filter (pmid %in% pmid [max (ct) > 24.5], .by = pmid) %>%

    rename (time = ct) %>%

    {.}

save (data_alfam2, file = "data_alfam2.Rdata")
```

# data_ml

```{r}
data_ml = data_alfam2 %>%

    mutate_if (is.character, as.factor) %>%

    mutate (t.incorp = ifelse (t.incorp <= 0.25, 0, t.incorp)) %>%

    mutate (t.incorp = ifelse (t.incorp == 0 & incorp == "none", 1000, t.incorp)) %>%
    mutate (t.incorp = replace_na (t.incorp, 1000)) %>%

    mutate (man.ph = replace_na (man.ph, median (man.ph, na.rm = TRUE)), .by = man.source) %>%

    # we round dt and time in order to compute the new climatic variables
    # cf the next chunk
    mutate (dt = round (dt, digits = 1)) %>%
    mutate (time = cumsum (dt), .by = pmid) %>%

    {.}
```

```{r}
# data = data_ml
# pmids = data %>% unique_vector (pmid)
# 
# data_ml = NULL
# 
# for (i in pmids) {
#         
#     data_tmp = data %>% filter (pmid == i)
#     
#     time_max = max (data_tmp$time)
#     
#     time = seq (0, (time_max - 0.1), by = 0.1)
#     
#     group = c(
#         sapply (c (1 : 6), function (j) rep (j, 40)) %>% as.vector,
#         rep (7, (length (time) - 240))
#     )
#     
#     # here we added 1+e-8 for numerical reasons, ('19.8 - 16.7 <= 3.1' return FALSE)
#     temp = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (air.temp))
#     wind = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (wind.2m))
#     rain = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (rain.rate))
# 
# 
#     df_temp = data.frame (time = time, temp = temp, group = group) %>% 
# 
#         summarise (temp = mean (temp), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = temp, names_prefix = "temp_")
#                    
# 
#     df_wind = data.frame (time = time, wind = wind, group = group) %>% 
# 
#         summarise (wind = mean (wind), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = wind, names_prefix = "wind_")
#                    
# 
#     df_rain = data.frame (time = time, rain = rain, group = group) %>% 
# 
#         summarise (rain = mean (rain), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = rain, names_prefix = "rain_")
# 
# 
#     data_tmp = cbind (
#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
#         df_temp,
#         df_wind,
#         df_rain
#     )
#                    
#     data_ml = rbind (data_ml, data_tmp)
#                                                          
# }

# save (data_ml, file = "data_ml.Rdata")

load (file = "data_ml.Rdata")
```

## data_random_forest

```{r}
data_random_forest = data_ml %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3)))

save (data_random_forest, file = "data_random_forest.Rdata")
```

## data_xgboost

```{r}
data_xgboost = data_ml %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3)))

save (data_xgboost, file = "data_xgboost.Rdata")
```

## data_lasso

```{r}
data_lasso = data_ml %>%

    # making the dummy variables --------
    mutate (man.source.pig = ifelse (man.source == "pig", 1, 0)) %>%
    select (- man.source) %>%

    mutate (app.mthd.bc = ifelse (app.mthd == "bc", 1, 0)) %>%
    mutate (app.mthd.ts = ifelse (app.mthd == "ts", 1, 0)) %>%
    mutate (app.mthd.os = ifelse (app.mthd == "os", 1, 0)) %>%
    select (- app.mthd) %>%

    mutate (incorp.shallow = ifelse (incorp == "shallow", 1, 0)) %>%
    mutate (incorp.deep = ifelse (incorp == "deep", 1, 0)) %>%
    select (- incorp) %>%
    # ---------

    mutate (e.cum = log (e.cum)) %>%

    {.}

save (data_lasso, file = "data_lasso.Rdata")
```

# data_scenarios

```{r}
data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%

    apply (2, quantile, probs = seq (0.01, 1, by = 0.01)) %>%
    as.data.frame %>%
    mutate (quantile = row_number () / 100) %>%
    pivot_longer (cols = - c (quantile)) %>%

    filter (quantile > 0.75) %>%
    ggplot () +
        geom_line (aes (x = quantile, y = value, color = name)) +
        scale_color_manual (values = Dark2)

```

```{r}
quartiles_climatic_variable = cbind (

    data_ml %>%

    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7) %>%

    apply (2, quantile, probs = c (0.25, 0.75)) %>%
    as.data.frame,

    data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%

    apply (2, quantile, probs = c (0.25, 0.95)) %>%
    as.data.frame
)
```

```{r}
grid_climatic_variable = quartiles_climatic_variable [, c(1 : 7)] %>% 
    cross_join (quartiles_climatic_variable [, c(8 : 14)]) %>% 
    cross_join (quartiles_climatic_variable [, c(15 : 21)]) %>%

    mutate (group_temp = ifelse (temp_1 == quartiles_climatic_variable[1, 1], "q1", "q2")) %>%
    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 8], "q1", "q2")) %>%
    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 16], "q1", "q2"))
```

```{r}
quartiles_plot_level_variable = data_ml %>%

    select (tan.app, app.rate, man.dm) %>%

    apply (2, quantile, probs = c (0.25, 0.75)) %>%
    as.data.frame
```

```{r}
grid_plot_level_variable = expand_grid (
    time = 72,
    tan.app = quartiles_plot_level_variable$tan.app,
    app.rate = quartiles_plot_level_variable$app.rate,
    man.dm = quartiles_plot_level_variable$man.dm,
    man.ph = 7.5,
    man.source = c ("pig", "cat"),
    incorp = c("none", "shallow"),
    app.mthd = c ("bc", "bsth", "os", "ts"),
    t.incorp = c (0, 1000)

) %>%

    filter (! (incorp == "none" & t.incorp %in% c(0))) %>%
    filter (! (incorp == "shallow" & t.incorp == 1000)) %>%
    filter (! (app.mthd %in% c ("bsth", "os", "ts") & incorp == "shallow"))
```

```{r}
nrow (grid_plot_level_variable)
# in the following calcul, the terms in parenthesis correspond to the variables time, tan.app, app.rate, man.dm, man.ph and man.source
# the last terms correspond to the variables incorp, app.mthd, and t.incorp

# incorp = none
(1 * 2 * 2 * 2 * 1 * 2) * 1 * 4 * 1 +

# incorp = shallow
(1 * 2 * 2 * 2 * 1 * 2) * 1 * 1 * 1
```

```{r}
data_scenarios = cross_join (grid_climatic_variable, grid_plot_level_variable) %>%

    mutate (pmid = row_number ()) %>%

    # reordering the variables
    select (pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp,
            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7,
            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7,
            group_temp, group_wind, group_rain)

save (data_scenarios, file = "data_scenarios.Rdata")
```

## data_scenarios_random_forest
```{r}
data_scenarios_random_forest = data_scenarios %>% 
    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3))) %>%

    {.}

save (data_scenarios_random_forest, file = "data_scenarios_random_forest.Rdata")
```

## data_scenarios_xgboost

```{r}
data_scenarios_xgboost = data_scenarios %>%

    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3))) %>%

    {.}

save (data_scenarios_xgboost, file = "data_scenarios_xgboost.Rdata")
```

## data_scenarios_lasso

```{r}
data_scenarios_lasso = data_scenarios %>%

    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    # making the dummy variables --------
    mutate (man.source.pig = ifelse (man.source == "pig", 1, 0)) %>%
    select (- man.source) %>%

    mutate (app.mthd.bc = ifelse (app.mthd == "bc", 1, 0)) %>%
    mutate (app.mthd.ts = ifelse (app.mthd == "ts", 1, 0)) %>%
    mutate (app.mthd.os = ifelse (app.mthd == "os", 1, 0)) %>%
    select (- app.mthd) %>%

    mutate (incorp.shallow = ifelse (incorp == "shallow", 1, 0)) %>%
    mutate (incorp.deep = ifelse (incorp == "deep", 1, 0)) %>%
    select (- incorp)
    # ---------

save (data_scenarios_lasso, file = "data_scenarios_lasso.Rdata")
```

## data_scenarios_alfam2

```{r}
data_scenarios_alfam2 = data_scenarios %>%

    mutate (t.incorp = na_if (t.incorp, 1000)) %>%

    select (- time) %>%

    pivot_longer (cols = c (10 : 30), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%

    select (- group_temp, - group_wind, - group_rain) %>%

    rename (air.temp = temp, wind.2m = wind, rain.rate = rain) %>%

    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "24", "7" = "72"))) %>%
    mutate (dt = ifelse (time == 72, (72 - 24), 4), .after = time) %>%

    mutate (rain.cum = cumsum (rain.rate), .by = pmid)

save (data_scenarios_alfam2, file = "data_scenarios_alfam2.Rdata")
```

### 2024-04-20 manuscript.qmd
-

@@
 ```{r}
-library ("ggplot2") 
-library ("gridExtra")
-library ("dplyr")       
-library ("paletteer")       
-library ("tibble") # rownames_to_column
-library ("tidyr") # pivot_longer
-library ("ALFAM2")
-library ("randomForest")
-library ("xgboost")
-library ("glmnet")
-library ("shapviz")
-library ("treeshap")
-
-set.seed (123)
+library ("ggplot2") # version 3.4.4
+library ("gridExtra") # version 2.3
+library ("dplyr") # version 1.1.2      
+library ("paletteer") # version 1.4.0
+library ("tibble") # version 3.2.1
+library ("tidyr") # version 1.3.0
+library ("ALFAM2") # version 3.7
+library ("randomForest") # version 4.7.1.1
+library ("xgboost") # version 1.7.6.1
+library ("glmnet") # version 4.1.8
+library ("shapviz") # version 0.9.3
+library ("treeshap") # version 0.3.1
 ```
@@
 ```{r}
-load (file = "scripts/data_alfam2.Rdata")
+load (file = "scripts/processed_data/data_alfam2.Rdata")
 ```
 
-```{r}
-plot = data_alfam2 %>%
+```{r, fig.width = 10, fig.height = 8}
+plot_data_description = data_alfam2 %>%
 
@@
 
+plot_data_description
+```
+
+```{r, include = FALSE}
 png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
-plot
+plot_data_description
 dev.off()
 
-plot
+```
+
+```{r}
+data_alfam2 %>%
+
+    select (air.temp, wind.2m, rain.rate, rain.cum) %>%
+
+    summary 
+```
+
+```{r}
+data_alfam2 %>%
+
+    filter (time == max (time), .by = pmid) %>%
+
+    select (e.cum, time, tan.app, app.rate, man.dm, man.ph, t.incorp) %>%
+
+    summary 
 ```
@@
 ```{r}
+# We use alfam2pars01 parameter without pH, like in Hafner et al, 2019
 pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]
@@
 
-        # Important to use alfam2pars01 parameters, which correspond to the parameters used in the paper
-        # New parameters are available, but leads to ohter results (less good results)
         pars = pars,
@@
 ```{r}
-load (file = "scripts/data_random_forest.Rdata")
+load (file = "scripts/processed_data/data_random_forest.Rdata")
 ```
@@
 
+set.seed (123)
 random_forest_model = randomForest (
@@
 
-        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8
+        mtry = 19, nodesize = 3, replace = FALSE, sample_frac = 0.8
 
@@
 ```{r}
-load (file = "scripts/data_xgboost.Rdata")
+load (file = "scripts/processed_data/data_xgboost.Rdata")
 ```
@@
 ```{r}
-
+set.seed (123)
 xgboost_model = xgboost (
@@
 
-    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,
+    max.depth = 6, nrounds = 300, eta = 0.3, min_child_weight = 0.5, subsample = 0.8,
 
@@
 ```{r}
-load (file = "scripts/data_lasso.Rdata")
+load (file = "scripts/processed_data/data_lasso.Rdata")
 ```
@@
 ```{r}
+set.seed (123)
 lasso_model = cv.glmnet (
@@
 
-```{r}
-plot = predictions_of_all_methods %>%
+```{r, fig.width = 10, fig.height = 8}
+plot_observed_vs_predicted_values = predictions_of_all_methods %>%
 
@@
 
-png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
-plot
-dev.off()
-
-plot
+plot_observed_vs_predicted_values
 ```
 
-## Evaluation metrics
+```{r, include = FALSE}
+png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
+plot_observed_vs_predicted_values
+dev.off()
+```
+
+### Evaluation metrics
 
@@
 
-```{r}
-# Evaluation for 72h cum. emission --------
+```{r, fig.width = 16, fig.height = 5}
+# Evaluation for 72h cum. emission
 df_plot = df_evaluation_metrics %>%
@@
     ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
-# --------
+```
 
-# Evaluation for 72h relative cum. emission ----------
+```{r, fig.width = 16, fig.height = 5}
+# Evaluation for 72h relative cum. emission
 df_plot = df_evaluation_metrics %>%
@@
     ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
-# --------
 ```
@@
 ```{r}
+set.seed (123)
 random_forest_model = randomForest (
@@
 ```{r}
-unified_model_rf <- randomForest.unify(random_forest_model, data_random_forest_calibration %>% select (- e.cum))
+df_tmp = data_random_forest_calibration %>% select (- e.cum)
 
-treeshap_rf <- treeshap(unified_model_rf, data_random_forest_calibration %>% select (- e.cum))
+unified_model_rf <- randomForest.unify (random_forest_model, df_tmp)
 
-shapley_rf <- shapviz(treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
+treeshap_rf <- treeshap (unified_model_rf, df_tmp, verbose = FALSE)
+
+shapley_rf <- shapviz (treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
 ```
 
-```{r}
+```{r, fig.width = 10, fig.height = 8}
 plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
@@
 
-```{r}
+```{r, fig.width = 10, fig.height = 12}
 df_importance_rf = importance (random_forest_model) %>%
@@
 
-### Shapley
-
 ```{r}
+set.seed (123)
 xgboost_model = xgboost (
@@
 
+### Shapley
+
 ```{r}
@@
 
-unified_model_xgboost <- xgboost.unify(xgboost_model, as.matrix(df_tmp))
+unified_model_xgboost <- xgboost.unify (xgboost_model, as.matrix(df_tmp))
 
-treeshap_xgboost <- treeshap(unified_model_xgboost, df_tmp)
+treeshap_xgboost <- treeshap (unified_model_xgboost, df_tmp, verbose = FALSE)
 
-shapley_xgb <- shapviz(treeshap_xgboost, X = df_tmp)
+shapley_xgb <- shapviz (treeshap_xgboost, X = df_tmp)
 ```
 
-```{r}
+```{r, fig.width = 10, fig.height = 8}
 plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
@@
 
-```{r}
+```{r, fig.width = 10, fig.height = 12}
 df_importance_xgb = xgb.importance (model = xgboost_model) %>%
@@
 ```
+
+## Lasso
+
+```{r}
+set.seed (123)
+lasso_model = cv.glmnet (
+
+    x = data_lasso %>% 
+        select (- pmid, - e.cum, - e.rel, - dataset) %>%
+        as.matrix %>%
+        {.},
+
+    y = data_lasso %>% 
+        select (e.cum) %>% 
+        as.matrix %>%
+        {.},
+
+    alpha = 1
+
+)
+```
+
 # Part 3 - Testing scenarios
@@
 ```{r}
-load (file = "scripts/data_scenarios.Rdata")
+load (file = "scripts/processed_data/data_scenarios.Rdata")
 
@@
 ```{r}
-load (file = "scripts/data_scenarios_random_forest.Rdata")
+load (file = "scripts/processed_data/data_scenarios_random_forest.Rdata")
 
@@
 ```{r}
-load (file = "scripts/data_scenarios_xgboost.Rdata")
+load (file = "scripts/processed_data/data_scenarios_xgboost.Rdata")
 
@@
     mutate (e.cum_hat_xgboost = xgboost_predictions_scenarios_vector, .before = time) %>%
-    mutate (e.cum_hat_xgboost = ifelse (e.cum_hat_xgboost < 0, 0, e.cum_hat_xgboost)) %>%
 
@@
 ```{r}
-load (file = "scripts/data_scenarios_lasso.Rdata")
+load (file = "scripts/processed_data/data_scenarios_lasso.Rdata")
 
@@
 ```{r}
-load (file = "scripts/data_scenarios_alfam2.Rdata")
+load (file = "scripts/processed_data/data_scenarios_alfam2.Rdata")
 
@@
 
-```{r}
-plot = data_scenarios %>% 
+```{r, fig.width = 10, fig.height = 10}
+plot_scenario_predictions = data_scenarios %>% 
     select (app.mthd, incorp, t.incorp, man.ph,
@@
 
+plot_scenario_predictions
+```
+
+```{r, include = FALSE}
 png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
-plot
+plot_scenario_predictions
 dev.off()
-
-plot
 ```
@@
 plot_shap_rf_2 = plot_shap_rf +
-    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
-    xlim (c (- 20, 58)) +
+    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
+    xlim (c (- 18, 45)) +
     theme (legend.position = "none", 
@@
 plot_shap_xgboost_2 = plot_shap_xgboost +
-    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
-    xlim (c (- 20, 58)) + 
+    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
+    xlim (c (- 18, 45)) + 
     theme (legend.position = "none", 

### 2024-04-20 scripts/data_processing.qmd
-

@@
 
+In this file, we build several dataframes that are used in the manuscript.qmd file. The following figure summarizes how these dataframes are built.
+
+![](figures/code_explanations.png){width=500, fig-align="center"}
+
+From the files of the ALFAM2 database (available in a Zenodo repository at https://zenodo.org/records/8109634), we first construct data_alfam2, which corresponds to the dataset used in Hafner et al (2019).
+From data_alfam2, we construct data_ml by retaining only the last observation of each PMID.
+The dynamics of climatic variables are preserved through class variables.
+From data_ml, we create data_lasso, data_xgboost, and data_random_forest, all of which contain the same information but are formatted according to the packages used to build the respective models.
+Additionally, from data_ml, we generate data_scenarios, which correspond to various scenarios on which we test the models.
+These scenarios are based on quantiles derived from data_ml.
+From data_scenarios we create four dataframes containing the same information but formatted according to the packages used for the different models.
+
 ```{r}
@@
 library ("paletteer")       
-library ("tibble") # rownames_to_column
-library ("tidyr") # pivot_longer
+library ("tidyr")
 ```
 
-```{r}
+```{r, include = FALSE}
 axis_text_size = 24 
@@
 ```{r}
+# these two files come from the ALFAM2 Zenodo repository (https://zenodo.org/records/8109634)
 data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
@@
 
+# these two files come from the supplementary material of Hafner et al (2019)
 calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
@@
 
-    filter (ct > 0 & ct <= 78) %>%
+    filter (ct <= 78) %>%
     select (j.NH3, e.cum, e.rel, pmid, ct, dt, air.temp, wind.2m, rain.rate, rain.cum, dataset)  %>%
@@
 
-    # we keep only trials that last more than 24h, to be in the same settings as for the machine learning models
-    filter (pmid %in% pmid [max (ct) > 24.5], .by = pmid) %>%
+    rename (time = ct)
 
-    rename (time = ct) %>%
-
-    {.}
-
-save (data_alfam2, file = "data_alfam2.Rdata")
+save (data_alfam2, file = "processed_data/data_alfam2.Rdata")
 ```
@@
     # cf the next chunk
-    mutate (dt = round (dt, digits = 1)) %>%
+    mutate (dt = round (dt, digits = 2)) %>%
     mutate (time = cumsum (dt), .by = pmid) %>%
@@
 # data = data_ml
-# pmids = data %>% unique_vector (pmid)
+# pmids = data %>% pull (pmid) %>% unique
 # 
@@
 #     
-#     time = seq (0, (time_max - 0.1), by = 0.1)
+#     time = seq (0, (time_max - 0.01), by = 0.01)
 #     
 #     group = c(
-#         sapply (c (1 : 6), function (j) rep (j, 40)) %>% as.vector,
-#         rep (7, (length (time) - 240))
+#         sapply (c (1 : 5), function (j) rep (j, 400)) %>% as.vector,
+#         rep (6, (length (time) - 2000))
 #     )
@@
 
-# save (data_ml, file = "data_ml.Rdata")
+# save (processed_data/data_ml, file = "data_ml.Rdata")
 
-load (file = "data_ml.Rdata")
+load (file = "processed_data/data_ml.Rdata")
 ```
@@
 
-save (data_random_forest, file = "data_random_forest.Rdata")
+save (data_random_forest, file = "processed_data/data_random_forest.Rdata")
 ```
@@
 
-save (data_xgboost, file = "data_xgboost.Rdata")
+save (data_xgboost, file = "processed_data/data_xgboost.Rdata")
 ```
@@
 
-save (data_lasso, file = "data_lasso.Rdata")
+save (data_lasso, file = "processed_data/data_lasso.Rdata")
 ```
@@
 
-```{r}
+```{r, fig.width = 10, fig.height = 5}
 data_ml %>%
 
-    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%
+    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%
 
@@
     mutate (quantile = row_number () / 100) %>%
-    pivot_longer (cols = - c (quantile)) %>%
+    pivot_longer (cols = - c (quantile), names_to = "variable") %>%
 
@@
     ggplot () +
-        geom_line (aes (x = quantile, y = value, color = name)) +
+        geom_line (aes (x = quantile, y = value, color = variable)) +
         scale_color_manual (values = Dark2)
@@
 
-    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
-            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7) %>%
+    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
+            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6) %>%
 
@@
 
-    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%
+    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%
 
@@
 ```{r}
-grid_climatic_variable = quartiles_climatic_variable [, c(1 : 7)] %>% 
-    cross_join (quartiles_climatic_variable [, c(8 : 14)]) %>% 
-    cross_join (quartiles_climatic_variable [, c(15 : 21)]) %>%
+quartiles_climatic_variable
+```
+
+```{r}
+grid_climatic_variable = quartiles_climatic_variable [, c(1 : 6)] %>% 
+    cross_join (quartiles_climatic_variable [, c(7 : 12)]) %>% 
+    cross_join (quartiles_climatic_variable [, c(13 : 18)]) %>%
 
     mutate (group_temp = ifelse (temp_1 == quartiles_climatic_variable[1, 1], "q1", "q2")) %>%
-    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 8], "q1", "q2")) %>%
-    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 16], "q1", "q2"))
+    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 7], "q1", "q2")) %>%
+    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 13], "q1", "q2"))
 ```
@@
     as.data.frame
+```
+
+```{r}
+quartiles_plot_level_variable
 ```
@@
     select (pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp,
-            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
-            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7,
-            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7,
+            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
+            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6,
+            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6,
             group_temp, group_wind, group_rain)
 
-save (data_scenarios, file = "data_scenarios.Rdata")
+save (data_scenarios, file = "processed_data/data_scenarios.Rdata")
 ```
@@
 
-save (data_scenarios_random_forest, file = "data_scenarios_random_forest.Rdata")
+save (data_scenarios_random_forest, file = "processed_data/data_scenarios_random_forest.Rdata")
 ```
@@
 
-save (data_scenarios_xgboost, file = "data_scenarios_xgboost.Rdata")
+save (data_scenarios_xgboost, file = "processed_data/data_scenarios_xgboost.Rdata")
 ```
@@
 
-save (data_scenarios_lasso, file = "data_scenarios_lasso.Rdata")
+save (data_scenarios_lasso, file = "processed_data/data_scenarios_lasso.Rdata")
 ```
@@
 
-    pivot_longer (cols = c (10 : 30), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%
+    pivot_longer (cols = c (10 : 27), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%
 
@@
 
-    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "24", "7" = "72"))) %>%
-    mutate (dt = ifelse (time == 72, (72 - 24), 4), .after = time) %>%
+    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "72"))) %>%
+    mutate (dt = ifelse (time == 72, (72 - 20), 4), .after = time) %>%
 
@@
 
-save (data_scenarios_alfam2, file = "data_scenarios_alfam2.Rdata")
+save (data_scenarios_alfam2, file = "processed_data/data_scenarios_alfam2.Rdata")
 ```

### 2024-04-20 scripts/random_forest_cross_validation.qmd
-

---
title: "Random forest cross-validation"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library (dplyr)
library (randomForest)
```

```{r}
load (file = "processed_data/data_random_forest.Rdata")
```

```{r}
data_random_forest_calibration = data_random_forest %>% 
    filter (dataset == "Calibration subset") %>%
    select (- pmid, - e.rel, - dataset)
```

```{r}
# set.seed (123)

# for (i in c (1 : 490)){
# 
#     calibration_subset_train = data_random_forest_calibration[- i, ]
# 
#     calibration_subset_eval = data_random_forest_calibration[i, ]
# 
#     # create hyper_grid search ------
#     hyper_grid <- expand.grid (
# 
#         mtry = c (8 : 28),
#         nodesize = c(1, 3, 5),
#         replace = c(TRUE, FALSE),
#         samp.fraction = c(0.63, 0.8, 1),
#         prediction = NA,
#         truth = NA,
#         executation_time = NA,
#         row_index = i
#     ) %>%
# 
#         filter (! (samp.fraction == 1 & replace == FALSE)) %>%
# 
#         mutate (sampsize = floor (samp.fraction * nrow (calibration_subset_train)))
#     # ------
# 
#     for (j in seq_len(nrow(hyper_grid))) {
# 
#         t0 = Sys.time()
# 
#         # fit model for jth hyperparameter combination
#         fit <- randomForest(
#                 formula      = e.cum ~ ., 
#                 data         = calibration_subset_train, 
#                 mtry         = hyper_grid$mtry[j],
#                 nodesize     = hyper_grid$nodesize[j],
#                 replace      = hyper_grid$replace[j],
#                 sampsize     = hyper_grid$sampsize[j],
#                 seed         = 123
#               )
# 
# 
#         tf = Sys.time()
# 
#         prediction = predict (fit, newdata = calibration_subset_eval %>% select (- e.cum))
#         truth = calibration_subset_eval$e.cum
# 
#         hyper_grid$prediction[j] = prediction
#         hyper_grid$truth[j] = truth
#         hyper_grid$executation_time[j] = difftime (tf, t0)
# 
#     }
#     
#     df_res = hyper_grid
# 
#     save (df_res, file = paste ("random_forest_cross_validation_results/df_res", i, "Rdata", sep = "."))
# 
# }
```

```{r}
files <- dir ("random_forest_cross_validation_results", pattern = ".Rdata", full.names = TRUE)

res <- do.call (rbind, lapply(files, function(x) get(load(x))))

res %>%

    summarise (rmse = sqrt (mean ((prediction - truth) ^ 2)), 
               .by = c (mtry, nodesize, replace, samp.fraction)) %>%

    arrange (rmse) %>%

    head (n = 20) 
```

### 2024-04-20 scripts/xgboost_cross_validation.qmd
-

---
title: "xgboost cross-validation"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library (dplyr)
library (xgboost)
```

```{r}
load (file = "processed_data/data_xgboost.Rdata")
```

```{r}
calibration_subset = data_xgboost %>%
    filter (dataset == "Calibration subset") %>%
    select (- e.rel, - dataset, - pmid)
```

```{r}
# set.seed (123)

# for (i in c (1 : 490)){
# 
#     calibration_subset_train = calibration_subset[- i, ]
# 
#     calibration_subset_eval = calibration_subset[i, ]
# 
#     # create hyper_grid search ------
#     hyper_grid <- expand.grid (
# 
#     max.depth = c (4, 6, 10, 14),
#     nrounds = c (50, 100, 150, 300),
#     eta = c (0.1, 0.2, 0.3),
#     min_child_weight = c (0.5, 1, 2),
#     subsample = c (0.8, 1),
#     prediction = NA,
#     truth = NA,
#     executation_time = NA,
#     row_index = i
#     )
#     # ------
# 
#     for (j in seq_len(nrow(hyper_grid))) {
# 
#         t0 = Sys.time()
# 
#         # fit model for ith hyperparameter combination
#         fit <- xgboost (
# 
#             data = xgb.DMatrix (
# 
#                 data = calibration_subset_train %>%
#                     select (- e.cum) %>%
#                     as.matrix %>%
#                     {.},
# 
#                 label = calibration_subset_train %>%
#                     select (e.cum) %>%
#                     as.matrix %>%
#                     {.}
# 
#             ),
# 
#             max.depth = hyper_grid$max.depth[j],
# 
#             nrounds = hyper_grid$nrounds[j],
# 
#             min_child_weight = hyper_grid$min_child_weight[j],
# 
#             eta = hyper_grid$eta[j],
# 
#             subsample = hyper_grid$subsample[j],
# 
#             verbose = FALSE
# 
#         )
# 
#         tf = Sys.time()
# 
#         prediction = predict (fit, newdata = calibration_subset_eval %>% select (- e.cum) %>% as.matrix)
#         truth = calibration_subset_eval$e.cum
# 
#         hyper_grid$prediction[j] = prediction
#         hyper_grid$truth[j] = truth
#         hyper_grid$executation_time[j] = difftime (tf, t0)
# 
#     }
# 
#     df_res = hyper_grid
# 
#     save (df_res, file = paste ("xgboost_cross_validation_results/df_res", i, "Rdata", sep = "."))
# 
# }
```

```{r}
files <- dir ("xgboost_cross_validation_results", pattern = ".Rdata", full.names = TRUE)

res <- do.call (rbind, lapply(files, function(x) get(load(x))))

res %>%

    summarise (rmse = sqrt (mean ((prediction - truth) ^ 2)), 
               .by = c (max.depth, nrounds, eta, min_child_weight, subsample)) %>%

    arrange (rmse)  %>%

    head (n = 20) 
```

### 2024-04-26 manuscript.qmd
removing relative emission of the Part 1

@@
 
-        select (pmid, time, e, er) %>%
+        select (pmid, time, e) %>%
 
         mutate (truth_e = data_alfam2$e.cum) %>%
-        mutate (truth_er = data_alfam2$e.rel) %>%
 
@@
 
-        select (pmid, time, e, er, truth_e, truth_er, dataset) %>%
+        select (pmid, time, e, truth_e, dataset) %>%
 
@@
 
-    mutate (e.rel_hat = e.cum_hat / tan.app) %>%
-
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)    
+    select (pmid, time, e = e.cum_hat, truth_e = e.cum, dataset)    
 ```
@@
 
-    mutate (e.rel_hat = e.cum_hat / tan.app) %>%
-
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)
+    select (pmid, time, e = e.cum_hat, truth_e = e.cum, dataset)
 ```
@@
 
-    mutate (e.rel_hat = e.cum_hat / tan.app) %>%
-
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset) 
+    select (pmid, time, e = e.cum_hat, truth_e = e.cum, dataset) 
 ```
@@
 ```{r}
-df_evaluation_metrics = rbind (
+df_evaluation_metrics = predictions_of_all_methods %>%
 
-    predictions_of_all_methods %>% 
     select (prediction = e, truth = truth_e, dataset, method) %>%
-        mutate (name = "72h cum. emission"),
-
-    predictions_of_all_methods %>% 
-        select (prediction = er, truth = truth_er, dataset, method) %>% 
-        mutate (name = "72h relative cum. emission")
-
-) %>%
 
@@
 
-        .by = c (name, dataset, method)
+        .by = c (dataset, method)
 
@@
 
-    rename (response = name) %>%
     pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
@@
 
-    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(response, name, dataset))
+    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(name, dataset))
 ```
@@
 df_plot = df_evaluation_metrics %>%
-    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
-    select (- name) %>%
+    filter (dataset == "Evaluation subset") %>%
     pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
@@
 df_plot = df_evaluation_metrics %>%
-    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
-    select (- name) %>%
+    filter (dataset == "Calibration subset") %>%
     pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%

### 2024-04-29 manuscript.qmd
adding three chunks in order to check in which conditions xgboost predicted a positive efficacy for the reduction strategies.

@@
 library ("treeshap") # version 0.3.1
+library ("kableExtra") # version 1.3.4
 ```
@@
 
+```{r}
+data_scenarios %>%
+    filter (efficacy_xgboost > 0) %>%
+    select (efficacy_xgboost, tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+    kable ()
+```
+
+```{r}
+data_scenarios %>%
+    filter (efficacy_xgboost > 0) %>%
+    select (tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+    lapply (table)
+```
+
+```{r, fig.width = 10, fig.height = 5}
+data_scenarios %>%
+  filter (efficacy_xgboost > 0) %>%
+  select (efficacy_xgboost, tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+  ggplot () +
+    geom_histogram (aes (efficacy_xgboost)) +
+    facet_wrap (~ tan.app + app.rate)
+```
+
 # Figures

### 2024-05-22 manuscript.qmd
adding table with importance values for rf and xgb
adding the country variable to the dataset to plot the residuals like in Hafner et al (2019) compute the relative emissions and the rmse to compare to Peng Xu et al adding a figure to show in which case xgboost predicted positive efficacy adding a figure to show the difference of the distributions between the data in Peng Xu et al (2024) and the data of Hafner et al (2019) cosmetics

--- 2024-04-20 manuscript.qmd
@@
 ```{r}
+library ("readxl")
 library ("ggplot2") # version 3.4.4
@@
 library ("glmnet") # version 4.1.8
+library ("treeshap") # version 0.3.1
 library ("shapviz") # version 0.9.3
-library ("treeshap") # version 0.3.1
+library ("kableExtra") # version 1.3.4
 ```
@@
 ```{r}
-axis_text_size = 24 
-axis_title_size = 30
+axis_text_size = 22 
+axis_title_size = 26
 title_size = 26
@@
 
+```{r}
+data_alfam2 %>% summarise (n = n(), .by = dataset)
+```
+
 ```{r, fig.width = 10, fig.height = 8}
@@
         xlab ("Time (h)") +
-        theme (axis.title.y = element_text (margin = ggplot2::margin (r = 25))) +
+        theme (axis.title.y = element_text (margin = ggplot2::margin (r = 25)),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
         # scale_color_manual (values = Dark2) +
@@
 dev.off()
-
 ```
@@
 
-        dat = data_alfam2 %>% select (- j.NH3, - e.cum, - e.rel, - dataset),
+        dat = data_alfam2 %>% select (- j.NH3, - e.cum, - e.rel, - dataset, - country),
 
@@
 
-        mutate (dataset = data_alfam2$dataset) %>%
+        mutate (dataset = data_alfam2$dataset,
+                country = data_alfam2$country,
+                app.mthd = data_alfam2$app.mthd) %>%
 
-        select (pmid, time, e, er, truth_e, truth_er, dataset) %>%
+        select (pmid, time, e, er, truth_e, truth_er, dataset, country, app.mthd) %>%
 
         filter (time == max (time), .by = pmid)
+```
+
+```{r, fig.width = 10, fig.height = 12}
+alfam2_predictions %>%
+    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hose", "ts" = "Trailing shoe", "os" = "Open slot injection")) %>%
+    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hose", "Trailing shoe", "Open slot injection"))) %>%
+    mutate (residuals = truth_er - er) %>%
+    pivot_longer (cols = c ("er", "truth_er", "residuals")) %>%
+    mutate (name = recode (name, "er" = "prediction", "truth_er" = "truth")) %>%
+    mutate (name = factor (name, levels = c ("truth", "prediction", "residuals"))) %>%
+    ggplot () +
+        geom_boxplot (aes (x = country, y = value, fill = name)) +
+        geom_hline (yintercept = 0, linetype = 2) +
+        facet_wrap (~ app.mthd, ncol = 1) +
+        ylab ("Relative 72h cum. emission or residual") +
+        labs (fill = "") +
+        theme (legend.position = "bottom",
+               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)))
 ```
@@
     filter (dataset == "Calibration subset") %>%
-    select (- pmid, - e.rel, - dataset)
+    select (- pmid, - e.rel, - dataset, - country)
 
@@
                 newdata = data_random_forest %>% 
-                    select (- pmid, - e.cum, - e.rel, - dataset)
+                    select (- pmid, - e.cum, - e.rel, - dataset, - country)
             )
@@
 
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)    
+    mutate (app.mthd = recode (as.character (app.mthd), "1" = "bc", "2" = "ts", "3" = "os", "4" = "bsth")) %>%
+
+    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset, country, app.mthd)    
 ```
@@
             filter (dataset == "Calibration subset") %>%
-            select (- pmid, - e.cum, - e.rel, - dataset) %>%
+            select (- pmid, - e.cum, - e.rel, - dataset, - country) %>%
             as.matrix %>%
@@
                 newdata = data_xgboost %>% 
-                        select (- pmid, - e.cum, - e.rel, - dataset) %>% 
+                        select (- pmid, - e.cum, - e.rel, - dataset, - country) %>% 
                         as.matrix
@@
 
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)
+    mutate (app.mthd = recode (as.character (app.mthd), "1" = "bc", "2" = "ts", "3" = "os", "4" = "bsth")) %>%
+
+    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset, country, app.mthd)
 ```
@@
         filter (dataset == "Calibration subset") %>%
-        select (- pmid, - e.cum, - e.rel, - dataset) %>%
+        select (- pmid, - e.cum, - e.rel, - dataset, - country) %>%
         as.matrix %>%
@@
     data_lasso %>% 
-            select (- pmid, - e.cum, - e.rel, - dataset) %>% 
+            select (- pmid, - e.cum, - e.rel, - dataset, - country) %>% 
             as.matrix
@@
 
+df_join = data_alfam2 %>% select (pmid, app.mthd) %>% distinct
+
 lasso_predictions = data_lasso %>%
+
+    left_join (df_join, by = "pmid") %>%
 
@@
 
-    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset) 
+    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset, country, app.mthd) 
 ```
@@
 )
+```
+
+```{r}
+predictions_of_all_methods %>% head
+```
+
+### Residuals
+
+```{r, fig.height = 12, fig.width = 10}
+plot_residuals_complete_dataset = predictions_of_all_methods %>%
+    mutate (residuals = truth_er - er) %>%
+    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hose", "ts" = "Trailing shoe", "os" = "Open slot injection")) %>%
+    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hose", "Trailing shoe", "Open slot injection"))) %>%
+    ggplot () +
+        geom_boxplot (aes (x = country, y = residuals, fill = method)) +
+        geom_hline (yintercept = 0, linetype = 2) +
+        scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
+        facet_wrap (~ app.mthd, ncol = 1) +
+        ylab ("Residual") +
+        labs (fill = "") +
+        theme (legend.position = "bottom",
+               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)))
+
+plot_residuals_complete_dataset
+```
+
+```{r, fig.height = 12, fig.width = 10}
+plot_residuals_evaluation_subset = predictions_of_all_methods %>%
+    filter (dataset == "Evaluation subset") %>%
+    mutate (residuals = truth_er - er) %>%
+    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hose", "ts" = "Trailing shoe", "os" = "Open slot injection")) %>%
+    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hose", "Trailing shoe", "Open slot injection"))) %>%
+    ggplot () +
+        geom_boxplot (aes (x = country, y = residuals, fill = method), varwidth = TRUE) +
+        geom_hline (yintercept = 0, linetype = 2) +
+        scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
+        facet_wrap (~ app.mthd, ncol = 1) +
+        ylab ("Residual") +
+        labs (fill = "") +
+        theme (legend.position = "bottom",
+               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)))
+
+plot_residuals_evaluation_subset
+```
+
+```{r, include = FALSE}
+png (file = "figures/plot_residuals_complete_dataset.png", width = 1000, height = 1100)
+plot_residuals_complete_dataset
+dev.off()
+
+png (file = "figures/plot_residuals_evaluation_subset.png", width = 1000, height = 1100)
+plot_residuals_evaluation_subset
+dev.off()
 ```
@@
                axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
-               axis.title.x = element_text (margin = ggplot2::margin (t = 15, b = 20))) +
+               axis.title.x = element_text (margin = ggplot2::margin (t = 15, b = 20)),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
 
@@
         select (prediction = e, truth = truth_e, dataset, method) %>% 
-        mutate (name = "72h cum. emission"),
+        mutate (response = "72h cum. emission"),
 
@@
         select (prediction = er, truth = truth_er, dataset, method) %>% 
-        mutate (name = "72h relative cum. emission")
+        mutate (response = "72h relative cum. emission")
 
@@
     summarise (
+
+        MSE = mean ((prediction - truth) ^ 2),
+
+        RMSE = sqrt (mean ((prediction - truth) ^ 2)),
 
@@
 
-        .by = c (name, dataset, method)
+        .by = c (dataset, method, response)
 
@@
 
-df_evaluation_metrics
+df_evaluation_metrics %>%  arrange (response, dataset) %>% kable ()
 ```
@@
 
-    rename (response = name) %>%
-    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
-    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE"))) %>% 
-    filter (name %in% c ("MAE", "MBE")) %>%
+    filter (response == "72h cum. emission") %>%
+
+    select (- MSE, - RMSE, - Pearsons_r, - ME) %>%
+
+    pivot_longer (cols = c (MAE, MBE)) %>%
 
@@
 
-    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(response, name, dataset))
+    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(name, dataset)) %>%
+
+    kable ()
 ```
@@
 ```{r, fig.width = 16, fig.height = 5}
-# Evaluation for 72h cum. emission
+# Evaluation on evalutation subset
 df_plot = df_evaluation_metrics %>%
-    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
-    select (- name) %>%
+    filter (response == "72h cum. emission") %>%
+    select (- MSE, - RMSE) %>%
+    filter (dataset == "Evaluation subset") %>%
     pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
@@
     xlab ("") + ylab ("") + labs (fill = "") +
-    theme (legend.position = "none ", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
+    theme (legend.position = "none ", 
+           strip.text.y = element_text(angle = 0), 
+           axis.text.x = element_blank(), 
+           axis.ticks.x = element_blank(),
+           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
     scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
@@
 ```{r, fig.width = 16, fig.height = 5}
-# Evaluation for 72h relative cum. emission
+# Evaluation on calibration subset
 df_plot = df_evaluation_metrics %>%
-    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
-    select (- name) %>%
+    filter (response == "72h cum. emission") %>%
+    select (- MSE, - RMSE) %>%
+    filter (dataset == "Calibration subset") %>%
     pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
@@
     xlab ("") + ylab ("") + labs (fill = "") +
-    theme (legend.position = "bottom", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
+    theme (legend.position = "bottom", 
+           strip.text.y = element_text(angle = 0), 
+           axis.text.x = element_blank(), 
+           axis.ticks.x = element_blank(),
+           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
     scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
@@
 
-        data = data_random_forest %>% select (- pmid, - e.rel, - dataset),
+        data = data_random_forest %>% select (- pmid, - e.rel, - dataset, - country),
 
@@
 
-unified_model_rf <- randomForest.unify (random_forest_model, df_tmp)
+unified_model_rf <- treeshap::randomForest.unify (random_forest_model, df_tmp)
 
-treeshap_rf <- treeshap (unified_model_rf, df_tmp, verbose = FALSE)
+treeshap_rf <- treeshap::treeshap (unified_model_rf, df_tmp, verbose = FALSE)
 
@@
 ```{r, fig.width = 10, fig.height = 8}
-plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
-    scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
+plot_shap_rf = shapviz::sv_importance (shapley_rf, kind = "beeswarm", max_display = 10L) +
+    scale_colour_gradient (low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
     theme (
@@
 ```{r, fig.width = 10, fig.height = 12}
-df_importance_rf = importance (random_forest_model) %>%
+df_importance_rf = randomForest::importance (random_forest_model) %>%
     as.data.frame %>%
@@
         scale_y_discrete (limits = df_importance_rf %>% arrange (`%IncMSE`) %>% pull (Variable)) +
+        theme (strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
         xlab ("Importance") +
@@
 ```
+
+```{r}
+df_importance_rf %>%
+    arrange (desc (`%IncMSE`)) %>%
+    head (n = 10) %>% kable ()
+```
+
 ## Xgboost
@@
         data = data_xgboost %>% 
-            select (- pmid, - e.cum, - e.rel, - dataset) %>%
+            select (- pmid, - e.cum, - e.rel, - dataset, - country) %>%
             as.matrix %>%
@@
 ```{r}
-df_tmp = data_xgboost %>% select (- pmid, - e.cum, - e.rel, - dataset)
+df_tmp = data_xgboost %>% select (- pmid, - e.cum, - e.rel, - dataset, - country)
 
-unified_model_xgboost <- xgboost.unify (xgboost_model, as.matrix(df_tmp))
+unified_model_xgboost <- treeshap::xgboost.unify (xgboost_model, as.matrix(df_tmp))
 
-treeshap_xgboost <- treeshap (unified_model_xgboost, df_tmp, verbose = FALSE)
+treeshap_xgboost <- treeshap::treeshap (unified_model_xgboost, df_tmp, verbose = FALSE)
 
-shapley_xgb <- shapviz (treeshap_xgboost, X = df_tmp)
+shapley_xgb <- shapviz::shapviz (treeshap_xgboost, X = df_tmp)
 ```
@@
 ```{r, fig.width = 10, fig.height = 8}
-plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
-  scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
+plot_shap_xgboost = shapviz::sv_importance (shapley_xgb, kind = "beeswarm", max_display = 10L) +
+  scale_colour_gradient (low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
   theme (legend.title = element_text (size = 30, margin = ggplot2::margin (r = 50, b = 100)),
@@
 ```{r, fig.width = 10, fig.height = 12}
-df_importance_xgb = xgb.importance (model = xgboost_model) %>%
+df_importance_xgb = xgboost::xgb.importance (model = xgboost_model) %>%
     mutate (Gain = Gain / max (Gain)) %>%
@@
         scale_y_discrete (limits = df_importance_xgb %>% arrange (Gain) %>% pull (Feature)) +
+        theme (strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
         xlab ("Importance") + ylab ("") +
@@
 plot_importance_xgboost
+```
+
+```{r}
+df_importance_xgb %>%
+    arrange (desc (Gain)) %>%
+    head (n = 10) %>% kable ()
 ```
@@
     x = data_lasso %>% 
-        select (- pmid, - e.cum, - e.rel, - dataset) %>%
+        select (- pmid, - e.cum, - e.rel, - dataset, - country) %>%
         as.matrix %>%
@@
         theme (legend.position = "none",
-               axis.ticks.y = element_blank()) +
+               axis.ticks.y = element_blank(),
+               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
         facet_wrap (~ method, ncol = 1)
@@
 
+```{r}
+data_scenarios %>%
+    filter (efficacy_xgboost > 0) %>%
+    select (efficacy_xgboost, tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+    kable ()
+```
+
+```{r}
+data_scenarios %>%
+    filter (efficacy_xgboost > 0) %>%
+    select (tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+    lapply (table)
+```
+
+```{r, fig.width = 20, fig.height= 8}
+data_tmp = data_scenarios %>%
+    filter (efficacy_xgboost > 0) %>%
+    select (efficacy_xgboost, tan.app, app.mthd, app.rate, man.source, incorp, group_temp, group_wind, group_rain) %>%
+    mutate (group_rain = ifelse (group_rain == "q1", "low", "high")) %>%
+    mutate (tan.app = paste ("TAN = ", tan.app),
+            app.rate = paste ("Application rate = ", app.rate),
+            group_rain = paste ("Rain = ", group_rain))
+
+plot = ggplot (data_tmp) +
+    geom_histogram (aes (efficacy_xgboost)) +
+    facet_wrap (~ tan.app + app.rate + group_rain) +
+    xlab ("Efficacy") +
+    ylab ("Number of predictions") +
+    theme (axis.title.y = element_text (margin = ggplot2::margin (r = 20)),
+           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)),
+           strip.text = element_text (size = 20))
+
+plot
+```
+
+```{r, include = FALSE}
+png (file = "figures/Xgboost_positive_predictions_plot.png", width = 1200, height = 800)
+plot
+dev.off()
+```
+
+```{r}
+data_scenarios %>%
+  filter (! (app.mthd == "bc" & incorp == "none")) %>%
+  summarise (mean_rf = mean (efficacy_random_forest), 
+             mean_alfam2 = mean (efficacy_alfam2), 
+             mean_gbf = mean (efficacy_xgboost), 
+             mean_lasso = mean (efficacy_lasso), .by = c ('app.mthd')) %>%
+  kable()
+```
+
 # Figures
@@
 plot_shap_rf_2 = plot_shap_rf +
-    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
+    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "#000080", color = "white") +
     xlim (c (- 18, 45)) +
@@
 plot_shap_xgboost_2 = plot_shap_xgboost +
-    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
+    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "#000080", color = "white") +
     xlim (c (- 18, 45)) + 
@@
 # Supplementary material
+
+```{r}
+data = read_excel("data/Peng_Xu_et_al_2024/Supplementary_Table_3_use2775.xlsx", sheet = 2)
+```
+
+```{r}
+head (data) %>% select (- Title) %>% kable()
+```
+
+```{r, fig.height = 8, fig.width = 10}
+plot_temperature_comparison = rbind (
+    data %>% select (temp = Tem) %>% mutate (data = "Data from Peng Xu et al (2024)"),
+    data_alfam2 %>% select (temp = air.temp) %>% mutate (data = "Data used for the alfam2 model")
+) %>%
+
+    ggplot () +
+        geom_histogram (aes (x = temp)) +
+        facet_wrap (~ data, ncol = 1, scales = "free") +
+        theme (strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)))+
+        xlab ("Air temperature (°C)") +
+        ylab ("Number of observations")
+
+plot_temperature_comparison
+```
+
+```{r, include = FALSE}
+png (file = "figures/plot_temperature_comparison.png", width = 1000, height = 600)
+plot_temperature_comparison
+dev.off()
+```
+
+```{r}
+rbind (
+    data %>% select (temp = Tem) %>% mutate (data = "Data from Peng Xu et al (2024)"),
+    data_alfam2 %>% select (temp = air.temp) %>% mutate (data = "Data used for the alfam2 model")
+) %>%
+
+  summarise (mean_temp = mean (temp), sd_temp = sd (temp), .by = data)
+```

### 2024-05-22 scripts/data_processing.qmd
adding a chunk to get the number of observations and trials
adding the country variable to the dataset

@@
             filter (pmid %in% pmids$pmid) %>%
-            select (pmid, tan.app, app.method, app.rate, man.dm, man.ph, man.source, incorp, time.incorp) %>%
+            select (pmid, country, tan.app, app.method, app.rate, man.dm, man.ph, man.source, incorp, time.incorp) %>%
             rename (app.mthd = app.method, t.incorp = time.incorp),
@@
 save (data_alfam2, file = "processed_data/data_alfam2.Rdata")
+```
+
+```{r}
+df_join = data_alfam2 %>% select (pmid, country) %>% distinct
+```
+
+```{r}
+data_alfam2 %>%
+  summarise (n_obs = n(), n_trial = length (unique (pmid)))
 ```
@@
 #     data_tmp = cbind (
-#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
+#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, country, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
 #         df_temp,
@@
 
-# save (processed_data/data_ml, file = "data_ml.Rdata")
+# save (data_ml, file = "processed_data/data_ml.Rdata")
 
 load (file = "processed_data/data_ml.Rdata")
+```
+
+```{r}
+dim (data_ml)
 ```

### 2024-05-22 scripts/random_forest_cross_validation.qmd
modification to not use the country variable for the inference

@@
     filter (dataset == "Calibration subset") %>%
-    select (- pmid, - e.rel, - dataset)
+    select (- pmid, - e.rel, - dataset, - country)
 ```

### 2024-05-23 scripts/data_processing.qmd
commenting the file and minor corrections

@@
 
+Construction of the data_alfam2 dataframe.
+
 ```{r}
 # these two files come from the ALFAM2 Zenodo repository (https://zenodo.org/records/8109634)
-data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
-data_plot_level = read.csv (file = "../data/ALFAM2_plot.csv")
+data_interval_level = read.csv (file = "../data/Hafner_et_al_2019/ALFAM2_interval.csv")
+data_plot_level = read.csv (file = "../data/Hafner_et_al_2019/ALFAM2_plot.csv")
 
 # these two files come from the supplementary material of Hafner et al (2019)
-calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
-evaluation_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc2.csv")
+calibration_subset_pmids = read.csv (file = "../data/Hafner_et_al_2019/1-s2.0-S1352231018308069-mmc1.csv")
+evaluation_subset_pmids = read.csv (file = "../data/Hafner_et_al_2019/1-s2.0-S1352231018308069-mmc2.csv")
 
@@
 # data_ml
+
+Construction of the dama_ml dataframe.
 
@@
 
+Construction of the data_random_forest dataframe.
+
 ```{r}
@@
 
+Construction of the data_xgboost dataframe.
+
 ```{r}
@@
 ## data_lasso
+
+Construction of the data_lasso dataframe.
 
@@
 # data_scenarios
+
+Construction of the data_scenarios dataframe.
 
@@
 ```{r}
-quartiles_climatic_variable
+quartiles_climatic_variable %>% round (digits = 1)
 ```
@@
 ```{r}
-quartiles_plot_level_variable
+quartiles_plot_level_variable %>% round (digits = 1)
 ```
@@
 ## data_scenarios_random_forest
+
+Construction of the data_scenarios_random_forest dataframe.
+
 ```{r}
@@
     mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
-    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3))) %>%
+    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2))) %>%
 
@@
 
+Construction of the data_scenarios_xgboost dataframe.
+
 ```{r}
@@
 ## data_scenarios_lasso
+
+Construction of the data_scenarios_lasso dataframe.
 
@@
     mutate (incorp.shallow = ifelse (incorp == "shallow", 1, 0)) %>%
-    mutate (incorp.deep = ifelse (incorp == "deep", 1, 0)) %>%
+    mutate (incorp.deep = 0) %>%
     select (- incorp)
@@
 
+Construction of the data_scenarios_alfam2 dataframe.
+
 ```{r}
@@
 
-    pivot_longer (cols = c (10 : 27), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%
+    pivot_longer (cols = contains ("temp_") | contains ("wind_") | contains ("rain_"), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%
 

### 2024-05-23 manuscript.qmd
commenting and minor corrections

@@
 ```{r}
-library ("readxl")
+library ("readxl") # version 1.4.3
 library ("ggplot2") # version 3.4.4
@@
 
-```{r}
+```{r, include = FALSE}
 axis_text_size = 22 
@@
 
-## Alfam2
+## ALFAM2
+
+Predictions of emissions using the ALFAM2 model.
 
@@
 
-```{r, fig.width = 10, fig.height = 12}
-alfam2_predictions %>%
-    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hose", "ts" = "Trailing shoe", "os" = "Open slot injection")) %>%
-    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hose", "Trailing shoe", "Open slot injection"))) %>%
-    mutate (residuals = truth_er - er) %>%
-    pivot_longer (cols = c ("er", "truth_er", "residuals")) %>%
-    mutate (name = recode (name, "er" = "prediction", "truth_er" = "truth")) %>%
-    mutate (name = factor (name, levels = c ("truth", "prediction", "residuals"))) %>%
-    ggplot () +
-        geom_boxplot (aes (x = country, y = value, fill = name)) +
-        geom_hline (yintercept = 0, linetype = 2) +
-        facet_wrap (~ app.mthd, ncol = 1) +
-        ylab ("Relative 72h cum. emission or residual") +
-        labs (fill = "") +
-        theme (legend.position = "bottom",
-               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
-               strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)))
-```
+## Random forest
 
-## Random forest
+Predictions of emissions using random forest.
 
@@
 ## Xgboost
+
+Predictions of emissions using gradient boosting.
 
@@
 ## Lasso
+
+Predictions of emissions using Lasso regression.
 
@@
 ```{r}
-predictions_of_all_methods %>% head
+predictions_of_all_methods %>% head %>% mutate_if (is.numeric, round, digits = 1) %>% kable
 ```
@@
 ### Residuals
+
+Comparison of residual for the calibration and evaluation subsets.
 
@@
 ```
+
+Comparison of residual for the evaluation subset.
 
@@
 
-    mutate_if (is.numeric, round, digits = 3)
+    mutate_if (is.numeric, round, digits = 2)
 
@@
     mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(name, dataset)) %>%
+    mutate (variation = round (variation, digits = 1)) %>%
+    mutate (value = round (value, digits = 1)) %>%
 
@@
     arrange (desc (`%IncMSE`)) %>%
-    head (n = 10) %>% kable ()
+    mutate_if (is.numeric, round, digits = 2) %>%
+    kable ()
 ```
@@
     arrange (desc (Gain)) %>%
-    head (n = 10) %>% kable ()
+    mutate_if (is.numeric, round, digits = 2) %>%
+    kable ()
 ```
@@
 
-## Alfam2
+## ALFAM2
 
@@
 
+Scenarios for which xgboost predicted a positive efficacy:
+
 ```{r}
@@
            strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)),
-           strip.text = element_text (size = 20))
+           strip.text = element_text (size = 24))
 
@@
              mean_lasso = mean (efficacy_lasso), .by = c ('app.mthd')) %>%
+  mutate_if (is.numeric, round, digits = 1) %>%
   kable()

### 2024-06-16 manuscript.qmd
minor corrections, adding unit in plots, computing median efficacy, ...

@@
   - "Céline Decuq"
-  - "Jean-Marc Giliot"
   - "David Makowski"
@@
         geom_line (aes (x = time, y = e.rel, group = pmid), linewidth = 0.15) +
-        ylab ("Cumul of relative emission") +
+        ylab ("Cumulative emission (frac. applied TAN)") +
         xlab ("Time (h)") +
@@
         facet_wrap (~ app.mthd, ncol = 1) +
-        ylab ("Residual") +
+        ylab ("Residual (frac. applied TAN)") +
         labs (fill = "") +
@@
 
-        xlab ("Observed values") + ylab ("Predicted values") +
+        xlab ("Observed values (kg/ha)") + ylab ("Predicted values (kg/ha)") +
 
@@
     mutate (group_rain = ifelse (group_rain == "q1", "low", "high")) %>%
-    mutate (tan.app = paste ("TAN = ", tan.app),
-            app.rate = paste ("Application rate = ", app.rate),
+    mutate (tan.app = paste ("TAN = ", round (tan.app, digits = 1), " kg/ha", sep = ""),
+            app.rate = paste ("Application rate = ", round (app.rate, digits = 1), " t/ha or m3/ha", sep = ""),
             group_rain = paste ("Rain = ", group_rain))
@@
     ylab ("Number of predictions") +
-    theme (axis.title.y = element_text (margin = ggplot2::margin (r = 20)),
+    theme (axis.title.y = element_text (margin = ggplot2::margin (r = 20), size = 18),
            strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0)),
-           strip.text = element_text (size = 24))
+           strip.text = element_text (size = 18),
+           axis.text = element_text (size = 18),
+           axis.title.x = element_text (size = 18))
 
@@
              mean_lasso = mean (efficacy_lasso), .by = c ('app.mthd')) %>%
+  mutate_if (is.numeric, round, digits = 1) %>%
+  kable()
+```
+
+```{r}
+data_scenarios %>%
+  filter (! (app.mthd == "bc" & incorp == "none")) %>%
+  summarise (median_rf = median (efficacy_random_forest), 
+             median_alfam2 = median (efficacy_alfam2), 
+             median_gbf = median (efficacy_xgboost), 
+             median_lasso = median (efficacy_lasso), .by = c ('app.mthd')) %>%
   mutate_if (is.numeric, round, digits = 1) %>%

### 2024-06-26 manuscript.qmd
updating figures, corrections of labels, ...

@@
     mutate (residuals = truth_er - er) %>%
-    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hose", "ts" = "Trailing shoe", "os" = "Open slot injection")) %>%
-    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hose", "Trailing shoe", "Open slot injection"))) %>%
+    mutate (app.mthd = recode (app.mthd, "bc" = "Broadcast", "bsth" = "Trailing hoses", "ts" = "Trailing shoes", "os" = "Open slot injection")) %>%
+    mutate (app.mthd = factor (app.mthd, levels = c ("Broadcast", "Trailing hoses", "Trailing shoes", "Open slot injection"))) %>%
     ggplot () +
@@
                             "bc shallow 0" = "incorporation",
-                            "bsth none -" = "trailing hose",
+                            "bsth none -" = "trailing hoses",
                             "os none -" = "open slot",
-                            "ts none -" = "trailing shoe")) %>%
+                            "ts none -" = "trailing shoes")) %>%
 
@@
         scale_fill_manual (values = Dark2[c(2 : 5)]) +
+        scale_y_discrete (limits = c ("open slot", "trailing shoes", "trailing hoses", "incorporation")) +
         ylab ("") + xlab ("Efficacy") +

### 2024-07-02 manuscript.qmd
renaming figures with number, and other minor changes

@@
 ```{r, fig.width = 10, fig.height = 8}
-plot_data_description = data_alfam2 %>%
+Figure_1 = data_alfam2 %>%
 
@@
 
-plot_data_description
+Figure_1
 ```
@@
 ```{r, include = FALSE}
-png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
-plot_data_description
+tiff (file = "figures/Figure_1.tiff", width = 1000, height = 700)
+Figure_1
 dev.off()
@@
 ```{r, include = FALSE}
-png (file = "figures/plot_residuals_complete_dataset.png", width = 1000, height = 1100)
+tiff (file = "figures/Figure_10.tiff", width = 1000, height = 1100)
 plot_residuals_complete_dataset
@@
 
-png (file = "figures/plot_residuals_evaluation_subset.png", width = 1000, height = 1100)
+tiff (file = "figures/Figure_11.tiff", width = 1000, height = 1100)
 plot_residuals_evaluation_subset
@@
 ```{r, fig.width = 10, fig.height = 8}
-plot_observed_vs_predicted_values = predictions_of_all_methods %>%
+Figure_2 = predictions_of_all_methods %>%
 
@@
 
-plot_observed_vs_predicted_values
+Figure_2
 ```
@@
 ```{r, include = FALSE}
-png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
-plot_observed_vs_predicted_values
+tiff (file = "figures/Figure_2.tiff", width = 1000, height = 700)
+Figure_2
 dev.off()
@@
 # Evaluation on evalutation subset
-df_plot = df_evaluation_metrics %>%
+df_plot_eval = df_evaluation_metrics %>%
     filter (response == "72h cum. emission") %>%
+    mutate (method = recode (method, "alfam2" = "ALFAM2", "xgboost" = "Gradient boosting", "random forest" = "Random forest", "lasso" = "Lasso")) %>%
+    mutate (method = factor (method, levels = c ("ALFAM2", "Lasso", "Random forest", "Gradient boosting"))) %>%
     select (- MSE, - RMSE) %>%
@@
 
-suppressWarnings(ggplot (df_plot) +
+plot_eval = suppressWarnings(ggplot (df_plot_eval) +
     geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
@@
            axis.text.x = element_blank(), 
+           axis.text.y = element_text (size = 18),
            axis.ticks.x = element_blank(),
-           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
+           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0), size = 24)) +
     scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
-    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
+    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 26)))
+
+plot_eval
+```
+
+```{r, include = FALSE}
+tiff (file = "figures/plot_eval.tiff", width = 1000, height = 320)
+plot_eval
+dev.off()
 ```
@@
 # Evaluation on calibration subset
-df_plot = df_evaluation_metrics %>%
+df_plot_calib = df_evaluation_metrics %>%
     filter (response == "72h cum. emission") %>%
+    mutate (method = recode (method, "alfam2" = "ALFAM2", "xgboost" = "Gradient boosting", "random forest" = "Random forest", "lasso" = "Lasso")) %>%
+    mutate (method = factor (method, levels = c ("ALFAM2", "Lasso", "Random forest", "Gradient boosting"))) %>%
     select (- MSE, - RMSE) %>%
@@
 
-suppressWarnings(ggplot (df_plot) +
+plot_calib = suppressWarnings(ggplot (df_plot_calib) +
     geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
@@
            axis.text.x = element_blank(), 
+           axis.text.y = element_text (size = 18),
            axis.ticks.x = element_blank(),
-           strip.text.x = element_text(margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0))) +
+           strip.text.x = element_text (margin = ggplot2::margin(t = 8, b = 8, r = 0, l = 0), size = 24),
+           axis.title.x = element_text (margin = ggplot2::margin (b = - 10))) +
     scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
-    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
+    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 26)))
+
+plot_calib
+```
+
+```{r, include = FALSE}
+tiff (file = "figures/plot_calib.tiff", width = 1000, height = 350)
+plot_calib
+dev.off()
 ```
@@
 ```{r, fig.width = 10, fig.height = 10}
-plot_scenario_predictions = data_scenarios %>% 
+Figure_6 = data_scenarios %>% 
     select (app.mthd, incorp, t.incorp, man.ph,
@@
 
-plot_scenario_predictions
+Figure_6
 ```
@@
 ```{r, include = FALSE}
-png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
-plot_scenario_predictions
+tiff (file = "figures/Figure_6.tiff", width = 1000, height = 800)
+Figure_6
 dev.off()
@@
 
-plot = ggplot (data_tmp) +
+Figure_8 = ggplot (data_tmp) +
     geom_histogram (aes (efficacy_xgboost)) +
@@
 
-plot
+Figure_8
 ```
@@
 ```{r, include = FALSE}
-png (file = "figures/Xgboost_positive_predictions_plot.png", width = 1200, height = 800)
-plot
+tiff (file = "figures/Figure_8.tiff", width = 1200, height = 800)
+Figure_8
 dev.off()
@@
 
-png (file = "figures/plot_shap.png", width = 1000, height = 900)
+tiff (file = "figures/Figure_5.tiff", width = 1000, height = 900)
 do.call ("grid.arrange", 
@@
 ```{r}
-png (file = "figures/plot_importance.png", width = 1300, height = 800)
+tiff (file = "figures/Figure_4.tiff", width = 1300, height = 800)
 grid.arrange (plot_importance_rf, plot_importance_xgboost, nrow = 1)
@@
 
+```{r, fig.height = 5, fig.width = 7}
+Figure_7 = data_alfam2 %>%
+    ggplot () +
+        geom_point (aes (x = man.ph, y = tan.app, color = man.source), size = 5) +
+        scale_color_manual (values = Dark2) +
+        xlab ("pH") +
+        ylab ("TAN") +
+        theme (legend.position = "bottom") +
+        labs (color = "Manure source") +
+        NULL
+
+Figure_7
+```
+
+```{r, include = FALSE}
+tiff (file = "figures/Figure_7.tiff", width = 1200, height = 800)
+Figure_7
+dev.off()
+```
+
 ```{r}
@@
 ```{r, fig.height = 8, fig.width = 10}
-plot_temperature_comparison = rbind (
+Figure_9 = rbind (
     data %>% select (temp = Tem) %>% mutate (data = "Data from Peng Xu et al (2024)"),
@@
 
-plot_temperature_comparison
+Figure_9
 ```
@@
 ```{r, include = FALSE}
-png (file = "figures/plot_temperature_comparison.png", width = 1000, height = 600)
-plot_temperature_comparison
+tiff (file = "figures/Figure_9.tiff", width = 1000, height = 600)
+Figure_9
 dev.off()

### 2024-09-11 manuscript.qmd
display of the lambda hyperparameter selected for lasso and removing paletteer dependancy

@@
 library ("dplyr") # version 1.1.2      
-library ("paletteer") # version 1.4.0
 library ("tibble") # version 3.2.1
@@
 
-Dark2 = paletteer_d("RColorBrewer::Dark2")
+Dark2 = c ("#1B9E77FF", "#D95F02FF", "#7570B3FF", "#E7298AFF", "#66A61EFF", "#E6AB02FF", "#A6761DFF", "#666666FF") 
 ```
@@
 ```{r}
+# lambda.1se is the largest value of lambda such that error is within 1 standard error of the 
+# lambda that minimizes the mean cross_validated error
+lasso_model$lambda.1se
+```
+
+```{r}
 lasso_predictions_vector = predict (
@@
             select (- pmid, - e.cum, - e.rel, - dataset, - country) %>% 
-            as.matrix
+            as.matrix,
+
+    s = "lambda.1se"
 )
@@
 
-    data_scenarios_lasso %>% as.matrix
+    data_scenarios_lasso %>% as.matrix,
+
+    s = "lambda.1se"
 )

### 2025-03-14 manuscript.qmd
Changing fonts in fugures

@@
     panel.background = element_rect(fill = "#f3faff"), panel.grid.major = element_line(colour = "white"), 
-    panel.spacing = unit(2, "lines")
+    panel.spacing = unit(2, "lines"),
+    text = element_text (family = "Times")
 )
@@
     mutate (strategy = paste (app.mthd, incorp, sep = " - ")) %>%
-    mutate (strategy = recode (strategy, "bc - none" = "A", "bc - shallow" = "B", "bc - deep" = "C", "bsth - none" = "D", "os - none" = "E", "ts - none" = "F")) %>%
+    #mutate (strategy = recode (strategy, "bc - none" = "A", "bc - shallow" = "B", "bc - deep" = "C", "bsth - none" = "D", "os - none" = "E", "ts - none" = "F")) %>%
+    mutate (strategy = recode (strategy, "bc - none" = "Broadcast", "bc - shallow" = "Shallow incorporation", "bc - deep" = "Deep incorporation", "bsth - none" = "Trailing hoses", "ts - none" = "Trailing shoes", "os - none" = "Open slot")) %>%
+
+    mutate (strategy = factor (strategy, levels = c ("Broadcast", "Shallow incorporation", "Deep incorporation", "Trailing hoses", "Trailing shoes", "Open slot"))) %>%
 
@@
 Figure_1
+
+png (file = "figures/Figure_1.png", width = 1000, height = 700)
+Figure_1
+dev.off()
 ```
@@
 
+```{r, include = FALSE}
+png (file = "figures/Figure_10.png", width = 1000, height = 1100)
+plot_residuals_complete_dataset
+dev.off()
+
+png (file = "figures/Figure_11.png", width = 1000, height = 1100)
+plot_residuals_evaluation_subset
+dev.off()
+```
+
 ### Observed vs predicted values
@@
 
-    mutate (method = recode (method, "alfam2" = "A", "lasso" = "B", "random forest" = "C", "xgboost" = "D")) %>%
+    mutate (method = recode (method, "alfam2" = "ALFAM2", "lasso" = "Lasso", "random forest" = "Random forest", "xgboost" = "Gradient boosting")) %>%
+
+    mutate (method = factor (method, levels = c ("ALFAM2", "Lasso", "Random forest", "Gradient boosting"))) %>%
 
@@
 
+```{r, include = FALSE}
+png (file = "figures/Figure_2.png", width = 1000, height = 700)
+Figure_2
+dev.off()
+```
+
 ### Evaluation metrics
@@
     mutate (value = round (value, digits = 1)) %>%
+
+    kable ()
+```
+
+```{r}
+df_evaluation_metrics %>%
+
+    filter (response == "72h cum. emission") %>%
+
+    select (- MSE, - RMSE, - MAE, - MBE) %>%
+
+    pivot_longer (cols = c (Pearsons_r, ME)) %>%
+
+    arrange (name, dataset) %>%
+
+    mutate (variation = (abs(value) / abs(value [method == "alfam2"]) - 1) * 100, .by = c(name, dataset)) %>%
+    mutate (variation = round (variation, digits = 1)) %>%
+    mutate (value = round (value, digits = 2)) %>%
 
@@
     mutate (`%IncMSE` = `%IncMSE` / max (`%IncMSE`)) %>%
-    mutate (type = "A")
+    mutate (type = "Random forest")
 
@@
     mutate (Gain = Gain / max (Gain)) %>%
-    mutate (type = "B")
+    mutate (type = "Gradient boosting")
 
@@
 tiff (file = "figures/Figure_6.tiff", width = 1000, height = 800)
+Figure_6
+dev.off()
+```
+
+```{r, include = FALSE}
+png (file = "figures/Figure_6.png", width = 1000, height = 800)
 Figure_6
@@
 
+```{r, include = FALSE}
+png (file = "figures/Figure_8.png", width = 1200, height = 800)
+Figure_8
+dev.off()
+```
+
 ```{r}
@@
 plot_shap_rf_2 = plot_shap_rf +
-    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "#000080", color = "white") +
+    annotate (geom = "label", label = "Random forest", size = 11, family = "Times",
+    label.padding = unit(0.8, "lines"), fontface = "bold", x = 38.5, y = 1.5, fill = "#000080", color = "white") +
     xlim (c (- 18, 45)) +
@@
 plot_shap_xgboost_2 = plot_shap_xgboost +
-    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "#000080", color = "white") +
+    annotate (geom = "label", label = "Gradient boosting", size = 11, family = "Times",
+    label.padding = unit(0.8, "lines"), fontface = "bold", x = 36.8, y = 1.5, fill = "#000080", color = "white") +
     xlim (c (- 18, 45)) + 
@@
 
+png (file = "figures/Figure_5.png", width = 1000, height = 900)
+do.call ("grid.arrange", 
+         c (list(plot_shap_rf_2, 
+                 plot_shap_xgboost_2, 
+                 legend), 
+            list(layout_matrix = as.matrix(c (1, 2, 3)),
+                  heights = c (0.8, 1, 0.2))))
+dev.off()
+
+```
+
+```{r}
 tiff (file = "figures/Figure_5.tiff", width = 1000, height = 900)
@@
 tiff (file = "figures/Figure_4.tiff", width = 1300, height = 800)
+grid.arrange (plot_importance_rf, plot_importance_xgboost, nrow = 1)
+dev.off()
+```
+
+```{r}
+png (file = "figures/Figure_4.png", width = 1300, height = 800)
 grid.arrange (plot_importance_rf, plot_importance_xgboost, nrow = 1)
@@
 
+```{r, fig.height = 5, fig.width = 7}
+Figure_7 = data_alfam2 %>%
+    mutate (man.source = recode (man.source, "cat" = "Cattle", "pig" = "Pig")) %>%
+    ggplot () +
+        geom_boxplot (aes (x = man.source, y = tan.app)) +
+        ylab ("TAN") +
+        xlab ("Manure source") +
+        theme (axis.text.x = element_text(margin = ggplot2::margin(t = 10))) +
+        NULL
+
+Figure_7
+```
+
 ```{r, include = FALSE}
@@
 
+```{r, include = FALSE}
+png (file = "figures/Figure_7.png", width = 1200, height = 800)
+Figure_7
+dev.off()
+```
+
 ```{r}
@@
 Figure_9 = rbind (
-    data %>% select (temp = Tem) %>% mutate (data = "Data from Peng Xu et al (2024)"),
+    data %>% select (temp = Tem) %>% mutate (data = "Data from Xu et al (2024)"),
     data_alfam2 %>% select (temp = air.temp) %>% mutate (data = "Data used for the alfam2 model")
@@
 
+```{r, include = FALSE}
+png (file = "figures/Figure_9.png", width = 1000, height = 600)
+Figure_9
+dev.off()
+```
+
 ```{r}
 rbind (
-    data %>% select (temp = Tem) %>% mutate (data = "Data from Peng Xu et al (2024)"),
+    data %>% select (temp = Tem) %>% mutate (data = "Data from Xu et al (2024)"),
     data_alfam2 %>% select (temp = air.temp) %>% mutate (data = "Data used for the alfam2 model")
